Portfolio Optimization Using Preference Relation Based on Statistical Arbitrage

### 2017-10-18 tex/presentation.tex
Final presentation.

  \begin{frame}
    \frametitle{Cons of statistical arbitrage method}
    \begin{itemize}
      \item low precision (in the sense of predicting price rise/fall): in most simulations \textbf{less than 50\%}
%      \item trading always done in pairs (\emph{short} and \emph{long} positions); requires possibility of \alert{short position} on the market
      \item doesn't expand well to more pairs
      \item frequent changes of the portfolio, resulting in \alert{significant transaction costs}
    \end{itemize}
  \end{frame}

  \section{Algorithm}
  \begin{frame}
    \frametitle{Preference relation}
    \begin{itemize}
      \item \textbf{binary relation:} $a \succ b$ --- $a$ is \emph{more preferred} than $b$
      \item \emph{incomparability} of entities: $a \sim b$ --- $a$ \emph{is not comparable} to $b$
      \item a more \emph{natural} way of comparing entities
      \item it is \emph{irreflexive}, \emph{asymmetric}, \emph{transitive}, and \emph{transitive of incomparability} 
    \end{itemize}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Simulation results --- subset from S\&P 500}
    \only<1>{
    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{graphics/result1.png}
      \caption*{Simulation results on a subset of 203 assets from S\&P 500 index.}
    \end{figure}}
    \only<2>{
      \begin{table}
        \centering
        \caption*{Simulation results on a subset of 203 assets from S\&P 500 index. Blue}
        \resizebox{\textwidth}{!}{\begin{tabular}{lrrr}
          \toprule
          \textbf{Method} & Buy \& Hold & Statistical arbitrage & Preference flow \\
          \midrule
          \textbf{Annual return} & 0.07622 & 0.63033 & \textbf{1.28000} \\
          \textbf{Volatility} & \textbf{0.15069} & 0.33532 & 0.78373 \\
          \textbf{Sharpe ratio} & 0.50582 & \textbf{1.87981} & 1.63322 \\
          \textbf{Average turnover rate} & / & 1.473211 & \textbf{0.55112} \\
          \midrule
          \textbf{Net profit with 0.10\% transaction costs} & 5.49572 & --2.74051 & 24.66204 \\
          \bottomrule
        \end{tabular}}
      \end{table}}
\end{frame}

@@

  \section{Conclusion}
  \begin{frame}
    \frametitle{Conclusion}
    \begin{itemize}
      \item \textbf{enhancement} of present methods of statistical arbitrage
      \item algorithm performs \textbf{better} when there is \textbf{more} assets available, allows for portfolio diversification
      \item algorithm works \textbf{well} even if \emph{short position} is not allowed
%      \item algorithm doesn't take \textbf{systemic risk} into account
    \end{itemize}
	\vfill
    \tiny{This work has been supported in part by the Croatian Science Foundation under the project 5349.}
    \includegraphics[height=1cm]{graphics/HRZZ.jpg}
  \end{frame}

\end{document}

>>>>

  \begin{frame}
    \frametitle{Cons of statistical arbitrage method}
    \begin{itemize}
      \item low precision (in the sense of predicting price rise/fall): in most simulations \textbf{less than 50\%}
%      \item trading always done in pairs (\emph{short} and \emph{long} positions); requires possibility of \alert{short position} on the market
      \item doesn't expand well to more pairs
      \item frequent changes of the portfolio result in \alert{significant transaction costs}
    \end{itemize}
  \end{frame}

  \section{Algorithm}
  \begin{frame}
    \frametitle{Preference relation}
    \begin{itemize}
      \item \textbf{binary relation:} $a \succ b$ %($a$ is \emph{more preferred} than $b$)
      \item \emph{incomparability} of entities: $a \sim b$ %($a$ \emph{is not comparable} to $b$)
      \item a more \emph{natural} way of comparing entities
      \item it is \emph{irreflexive}, \emph{asymmetric}, \emph{transitive}, and \emph{transitive in incomparability} 
    \end{itemize}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Simulation results --- subset from S\&P 500}
    \only<1>{
    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{graphics/result1.png}
      \caption*{Simulation results on a subset of 203 assets from S\&P 500 index.}
    \end{figure}}
    \only<2>{
      \begin{table}
        \centering
        \caption*{Simulation results on a subset of 203 assets from S\&P 500 index.}
        \resizebox{\textwidth}{!}{\begin{tabular}{lrrr}
          \toprule
          \textbf{Method} & Buy \& Hold & Statistical arbitrage & Preference flow \\
          \midrule
          \textbf{Annual return} & 0.07622 & 0.63033 & \textbf{1.28000} \\
          \textbf{Volatility} & \textbf{0.15069} & 0.33532 & 0.78373 \\
          \textbf{Sharpe ratio} & 0.50582 & \textbf{1.87981} & 1.63322 \\
          \textbf{Average turnover rate} & / & 1.473211 & \textbf{0.55112} \\
          \midrule
          \textbf{Net profit with 0.10\% transaction costs} & 5.49572 & --2.74051 & 24.66204 \\
          \bottomrule
        \end{tabular}}
      \end{table}}
\end{frame}

@@

  \section{Conclusion}
  {
  \setbeamertemplate{footline}{
  	\begin{center}
	  	\tiny{This work has been supported in part by the Croatian Science Foundation under the project 5349.} \\
  		\includegraphics[height=0.5cm]{graphics/HRZZ.jpg}
    \end{center}
  }
  \begin{frame}
    \frametitle{Conclusion}
    \begin{itemize}
      \item \textbf{enhancement} of present methods of statistical arbitrage
      \item algorithm performs \textbf{better} when there is \textbf{more} assets available, allows for portfolio diversification
      \item algorithm works \textbf{well} even if \emph{short position} is not allowed
%      \item algorithm doesn't take \textbf{systemic risk} into account
    \end{itemize}

  \end{frame}
  }
\end{document}
